Demo arXiv template

### 2022-11-27 paper/paper.qmd
more content in temporal processing

An overview of the pipeline is given in @fig-pipeline-steps to illustrate the construction from raw data to the final indexes.
The available intermediate steps include temporal processing, spatial aggregation, variable transformation, scaling, dimension reduction, normalising, benchmarking, and simplification. 
These steps are not required to be arranged in a fixed order and not all the steps have to appear when constructing each index. 
The pipeline includes modules for operations in the spatial, temporal, and multivariate aspect of the data, as well as modules for comparing and communicating indexes (benchmarking and simplification).
While the pipeline starts with the raw data, steps prior to this, i.e. define a useful concept to construct the index, collect high quality data relevant for measuring the concept, are also cricial to the sucess of an index.
The framework proposed in this paper assumes the objective of an index has been defined and relevant variables have been collected and quality controlled.

@@

Let $\mathbf{x}(\mathbf{s};\mathbf{t})$ denote the raw data with spatial, temporal, and multivariate aspect: the spatial dimension $\mathbf{s} =  (s_1, s_2, \cdots, s_n)^\prime$ is defined in the 2D space: $\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2$, the temporal dimension $\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime$ is defined in the 1D space: $\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}$. When more than one variable are invovled, the multivariate data can also be written as: $\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime$.

@@

While many variables are collected separately at each time period, the underlying phenomenon, characterised by the index, is continuous. 
For example, precipitation can be measured daily or monthly, a meteorological drought is characterised by a prelonged lack of precipitation. 
Temporal proocessing is a step in the pipeline that includes operations in the temporal dimension and it can be generally represented by $$f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),$$ where $\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}$ is the parameters associated with the temporal operation and $d_{\psi}$ is the number of parameter of $\psi$. The purpose of temporal aggregation can be to aggregate variable across time or to borrow information from nearby time periods to smooth or adjust the current value. The drought index SPI is constructed from the aggregated precipitation with a time scale parameter $\psi = k$. The aggregation step in SPI can be written as $x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j)$ where $j^\prime$ is the time dimension after the aggregation.

>>>>

An overview of the pipeline is given in @fig-pipeline-steps to illustrate the construction from raw data to the final indexes.
The pipeline includes modules for operations in the spatial, temporal, and multivariate aspect of the data: temporal processing, spatial aggregation, variable transformation, scaling, dimension reduction, and normalising, as well as modules for comparing and communicating indexes (benchmarking and simplification).
These steps are not required to be arranged in a fixed order and not all the steps have to appear when constructing each index. 
While the pipeline starts with the raw data, steps prior to this, i.e. define a useful concept to construct the index, collect high quality data relevant for measuring the concept, are also cricial to the sucess of an index.
The framework proposed in this paper assumes the objective of an index has been defined and relevant variables have been collected and quality controlled.

@@

Let $\mathbf{x}(\mathbf{s};\mathbf{t})$ denote the raw data with spatial, temporal, and multivariate aspect: the spatial dimension $\mathbf{s} =  (s_1, s_2, \cdots, s_n)^\prime$ is defined in the 2D space: $\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2$, the temporal dimension $\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime$ is defined in the 1D space: $\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}$. When more than one variable are involved, the multivariate data can also be written as: $\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime$.

@@

The construction of an index sometimes needs to consider information from neighbouring time periods.
Temporal processing is a general operator on the time dimension of the data in the form of  $$f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),$$ where $\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}$ is the parameters associated with the temporal operation and $d_{\psi}$ is the number of parameter of $\psi$. 
An typical example of temporal processing is aggregation in the drought index SPI to measure the lack of precipitation for meteorological drought.
In SPI, monthly precipitation is aggregated by a time scale parameter $k$: $x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),$ where $j^\prime$ is the new time index after the aggregation. In this notation, each spatial location is seaprately aggregated and precipitation is ummed from $k$ month back, $j^\prime - k + 1$, to the current, $j^\prime$ period to create the aggregated series, indexed by $j^\prime$.

The choice of time scales parameter $k$ can result in variation in the calculated index values: a small $k$ of 3 or 6 month produces the index more sensitive to individual months, while a large $k$ of 24 or 36, an equivalent to a 2- or 3-year aggregation, gives dryness information relative to the long term condition. As will be shown in section [SECTION EXAMPLE], this variation may even lead to conflicting conclusions on the dry/wet condition of the area, highlighting the importance to account for index uncertainty when interpreting index values for decision making.

### 2022-11-27 paper/paper.tex
more content in temporal processing

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

An overview of the pipeline is given in Figure~\ref{fig-pipeline-steps}
to illustrate the construction from raw data to the final indexes. The
available intermediate steps include temporal processing, spatial
aggregation, variable transformation, scaling, dimension reduction,
normalising, benchmarking, and simplification. These steps are not
required to be arranged in a fixed order and not all the steps have to
appear when constructing each index. The pipeline includes modules for
operations in the spatial, temporal, and multivariate aspect of the
data, as well as modules for comparing and communicating indexes
(benchmarking and simplification). While the pipeline starts with the
raw data, steps prior to this, i.e.~define a useful concept to construct
the index, collect high quality data relevant for measuring the concept,
are also cricial to the sucess of an index. The framework proposed in
this paper assumes the objective of an index has been defined and
relevant variables have been collected and quality controlled.

@@

Let \(\mathbf{x}(\mathbf{s};\mathbf{t})\) denote the raw data with
spatial, temporal, and multivariate aspect: the spatial dimension
\(\mathbf{s} = (s_1, s_2, \cdots, s_n)^\prime\) is defined in the 2D
space: \(\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2\), the
temporal dimension \(\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime\) is
defined in the 1D space:
\(\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}\). When more than
one variable are invovled, the multivariate data can also be written as:
\(\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime\).

@@

While many variables are collected separately at each time period, the
underlying phenomenon, characterised by the index, is continuous. For
example, precipitation can be measured daily or monthly, a
meteorological drought is characterised by a prelonged lack of
precipitation. Temporal proocessing is a step in the pipeline that
includes operations in the temporal dimension and it can be generally
represented by \[f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),\] where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\). The purpose of temporal aggregation can be to
aggregate variable across time or to borrow information from nearby time
periods to smooth or adjust the current value. The drought index SPI is
constructed from the aggregated precipitation with a time scale
parameter \(\psi = k\). The aggregation step in SPI can be written as
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j)\)
where \(j^\prime\) is the time dimension after the aggregation.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, enhanced, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

An overview of the pipeline is given in Figure~\ref{fig-pipeline-steps}
to illustrate the construction from raw data to the final indexes. The
pipeline includes modules for operations in the spatial, temporal, and
multivariate aspect of the data: temporal processing, spatial
aggregation, variable transformation, scaling, dimension reduction, and
normalising, as well as modules for comparing and communicating indexes
(benchmarking and simplification). These steps are not required to be
arranged in a fixed order and not all the steps have to appear when
constructing each index. While the pipeline starts with the raw data,
steps prior to this, i.e.~define a useful concept to construct the
index, collect high quality data relevant for measuring the concept, are
also cricial to the sucess of an index. The framework proposed in this
paper assumes the objective of an index has been defined and relevant
variables have been collected and quality controlled.

@@

Let \(\mathbf{x}(\mathbf{s};\mathbf{t})\) denote the raw data with
spatial, temporal, and multivariate aspect: the spatial dimension
\(\mathbf{s} = (s_1, s_2, \cdots, s_n)^\prime\) is defined in the 2D
space: \(\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2\), the
temporal dimension \(\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime\) is
defined in the 1D space:
\(\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}\). When more than
one variable are involved, the multivariate data can also be written as:
\(\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime\).

@@

The construction of an index sometimes needs to consider information
from neighbouring time periods. Temporal processing is a general
operator on the time dimension of the data in the form of
\[f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t})),\] where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\). An typical example of temporal processing is
aggregation in the drought index SPI to measure the lack of
precipitation for meteorological drought. In SPI, monthly precipitation
is aggregated by a time scale parameter \(k\):
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j),\)
where \(j^\prime\) is the new time index after the aggregation. In this
notation, each spatial location is seaprately aggregated and
precipitation is ummed from \(k\) month back, \(j^\prime - k + 1\), to
the current, \(j^\prime\) period to create the aggregated series,
indexed by \(j^\prime\).

The choice of time scales parameter \(k\) can result in variation in the
calculated index values: a small \(k\) of 3 or 6 month produces the
index more sensitive to individual months, while a large \(k\) of 24 or
36, an equivalent to a 2- or 3-year aggregation, gives dryness
information relative to the long term condition. As will be shown in
section {[}SECTION EXAMPLE{]}, this variation may even lead to
conflicting conclusions on the dry/wet condition of the area,
highlighting the importance to account for index uncertainty when
interpreting index values for decision making.
